---
title: "Sensitivity of primary productivity to climate is mediated by biodiversity"
author: "Brunno F Oliveira"
date: "as.character(Sys.Date())"
output:
  pdf_document: default
  html_document: default
---

*** 

\newpage

# Packages versions:
```{r info, message=FALSE,echo=F}
info <- sessionInfo()
```

We used `r info[1]$R$ver` and the following packages:

# Load packages
```{r packages, message=FALSE, include=FALSE}

rm(list=ls())
gc()

list.of.packages <- c("raster","maps","nlme","spdep","rgdal","viridis","ggplot2","rworldmap","gridExtra","cowplot","effects","PerformanceAnalytics","dplyr","Rphylopars","pbapply","parallel","doParallel","reshape","picante","FD","corrplot","ggvenn","WorldFlora")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

sapply(list.of.packages, require, character.only = TRUE)

```

# Load image
```{r load, message=FALSE, warning=FALSE, include=FALSE}

load("R/FixData.RData")

```

# Load coastal lines
```{r land, eval=FALSE, message=FALSE, warning=FALSE}

land <- readOGR("D:/GIS/Shp files/WE_contour.shp")
crs(land) <-"+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0 "

# project to equal area
land <- spTransform(land, CRS("+proj=moll +datum=WGS84"))

```

# Load occurrence data
Data downloaded from BIEN version 4.1
This comes at a 1/2 dg (~50km x 50km) resolution

```{r occr, eval=FALSE, message=FALSE, warning=FALSE}

plants <- readRDS("Data/BIEN_skinny_ranges.RDs")
plants <- data.frame(plants)
names(plants) <- c("sps","cell")

# get sp names
sp.ranges <- unique(plants$sps) 

# N species with range data
length(sp.ranges) # 98826

# calc richness with the full data
richness.full <- plants %>%
  group_by(cell) %>% 
  summarise(rich = length(unique(sps)))

rownames(richness.full) <- richness.full$cell

template_raster <- raster(xmn = -180, xmx = 180, 
                          ymn = -60, ymx = 85, 
                          resolution = .5, crs=crs(land))

cells <- richness.full$cell
XY <- data.frame(xyFromCell(template_raster, as.numeric(cells)))
names(XY) <- c("x","y")

plot(rasterFromXYZ(data.frame(XY, richness.full$rich)))

```

# Load phylogeny
We downloaded 100 random phylogenies from BIEN. Seed = 666.
```{r phy, eval=FALSE, message=FALSE, warning=FALSE}

phy <- readRDS("Data/BIEN_phylogeny_n100.RDS")

# get sp names
sp.phy <- unique(phy[[1]]$tip.label)
# N species in phy
length(sp.phy) # 81274

```

# Load trait data

## BIEN

We downloaded data for the 6 most complete traits from BIEN: 
* leaf area per leaf dry mass (SLA)
* whole plant height / maximum whole plant height (Height)
* seed mass (SeedMass)
* leaf nitrogen content per leaf dry mass (N)
* leaf phosphorus content per leaf dry mass (P)
* longest whole plant longevity (Longevity)
* stem wood density (WoodDens)

```{r traits_bien, eval=FALSE, message=FALSE, warning=FALSE}
# Load traits from BEIN
traitsBIEN <- readRDS("Data/BIEN_traits.Rds")
names(traitsBIEN) <- sapply(traitsBIEN, function(x) unique(x$trait_name))
# combine trait data into a single table
traitsBIEN <- do.call("rbind", traitsBIEN)
# get only columns of interest
traitsBIEN <- traitsBIEN[,1:3]
colnames(traitsBIEN) <- c("sps", "trait", "value")
# omit NAs
traitsBIEN <- na.omit(traitsBIEN)
# Fix plant height trait
traitsBIEN$trait[which(traitsBIEN$trait=="maximum whole plant height")] <- "whole plant height"
# Fix plant plant longevity
traitsBIEN$trait[which(traitsBIEN$trait=="longest whole plant longevity")] <- "maximum whole plant longevity"
# delete these two traits
traitsBIEN <- traitsBIEN[-which(traitsBIEN$trait=="whole plant growth form"),]
traitsBIEN <- traitsBIEN[-which(traitsBIEN$trait=="whole plant dispersal syndrome"),]
# trait value as numeric
traitsBIEN$value <- as.numeric(traitsBIEN$value)
# fix sp names
traitsBIEN$sps <- gsub(" ","_",traitsBIEN$sps)
# get sp names
sp.traitsBIEN <- unique(traitsBIEN$sps)
# N species in traitBIEN
length(sp.traitsBIEN) # 21629

# N species per trait
obs.BIEN <- traitsBIEN %>%
  group_by(trait) %>%
  summarise(N_sps = length(unique(sps)),
            N_obs = length(sps),
            Mean_obs_sps = round(N_obs/N_sps,2))

DT::datatable(obs.BIEN)

```
We downloaded trait data from BIEN, giving a total of `r length(sp.traitsBIEN)` species. 


## TRY
We downloaded data for the 9 most complete traits from TRY: 
* Leaf area per leaf dry mass (SLA)
* Leaf nitrogen content per leaf dry mass (N)
* Leaf phosphorus content per leaf dry mass (P)
* Plant height vegetative (Height)
* Plant lifespan (Longevity)
* Root rooting depth (RootDepth)
* Seed dry mass (SeedMass)
* tem specific density or wood density (WoodDens)

```{r traits_try, eval=FALSE, message=FALSE, warning=FALSE}
# Load traits from TRY
traitsTRY <- readRDS("Data/TRYdata_cont.RDS")
traitsTRY <- traitsTRY[,c("AccSpeciesName","TraitName","StdValue")]
colnames(traitsTRY) <- c("sps", "trait", "value")
traitsTRY$sps <- gsub(" ","_",traitsTRY$sps)
# Remove CN trait
traitsTRY <- traitsTRY[-which(traitsTRY$trait=="Leaf carbon/nitrogen (C/N) ratio"),]

# get sp names
sp.traitsTRY <- unique(traitsTRY$sps)
# N species in traitsTRY
length(sp.traitsTRY) # 51655

# N species per trait
obs.t <- traitsTRY %>%
  group_by(trait) %>%
  summarise(N_sps = length(unique(sps)),
            N_obs = length(sps),
            Mean_obs_sps = round(N_obs/N_sps,2))
DT::datatable(obs.t)

##########
# species in BIEN but not in TRY
length(which(!sp.traitsBIEN %in% sp.traitsTRY)) # 3078
# species in TRY but not in BIEN
length(which(!sp.traitsTRY %in% sp.traitsBIEN)) # 33104. PS: TRY data is global, while BIEN is only for the Americas
##########

```

We downloaded trait data from TRY, giving a total of `r length(sp.traitsTRY)` species. 


## Merge

```{r traits_merge, eval=FALSE, message=FALSE, warning=FALSE}
# merge both datasets
# fix names
colnames(traitsBIEN)
colnames(traitsTRY)

# fix trait names
unique(traitsBIEN$trait)
traitsBIEN$trait[which(traitsBIEN$trait=="whole plant height")] <- "Height"
traitsBIEN$trait[which(traitsBIEN$trait=="leaf nitrogen content per leaf dry mass")] <- "N"
traitsBIEN$trait[which(traitsBIEN$trait=="leaf area per leaf dry mass")] <- "SLA"
traitsBIEN$trait[which(traitsBIEN$trait=="seed mass")] <- "SeedMass"
traitsBIEN$trait[which(traitsBIEN$trait=="leaf phosphorus content per leaf dry mass")] <- "P"
traitsBIEN$trait[which(traitsBIEN$trait=="maximum whole plant longevity")] <- "Longevity"
traitsBIEN$trait[which(traitsBIEN$trait=="stem wood density")] <- "WoodDens"

unique(traitsTRY$trait)
traitsTRY$trait[which(traitsTRY$trait=="Plant height vegetative")] <- "Height"
traitsTRY$trait[which(traitsTRY$trait=="Plant lifespan (longevity)")] <- "Longevity"
traitsTRY$trait[which(traitsTRY$trait=="Leaf phosphorus (P) content per leaf dry mass")] <- "P"
traitsTRY$trait[which(traitsTRY$trait=="Leaf nitrogen (N) content per leaf dry mass")] <- "N"
traitsTRY$trait[which(traitsTRY$trait=="Leaf area per leaf dry mass (specific leaf area, SLA or 1/LMA): undefined if petiole is in- or excluded")] <- "SLA"
traitsTRY$trait[which(traitsTRY$trait=="Stem specific density (SSD) or wood density (stem dry mass per stem fresh volume)")] <- "WoodDens"
traitsTRY$trait[which(traitsTRY$trait=="Seed dry mass")] <- "SeedMass"
traitsTRY$trait[which(traitsTRY$trait=="Root rooting depth")] <- "RootDepth"

# combine
traitslong <- rbind(traitsBIEN,traitsTRY)
traitslong$sps <- gsub(" ","_",traitslong$sps)

# get median value per trait
traitslong.med <- traitslong %>%
  group_by(sps, trait) %>%
  summarise(trait_value = median(value, na.rm = T))
# from long to short
traits <- cast(traitslong.med, sps~trait)

# Any sp without traits?
any(which(rowSums(traits[,-1])==0))

# get sp names
sp.traits <- unique(traits$sps)
# N species with traits
length(sp.traits) # 54733

# The basic species list to work are those with range and phylogenetic data
# These were call sp.work
sp.work <- sp.phy[which(sp.phy %in% sp.ranges)]

# N species per trait
obs.total <- traitslong %>%
  group_by(trait) %>%
  summarise(N_sps = length(unique(sps)),
            N_obs = length(sps),
            Mean_obs_sps = round(N_obs/N_sps,2),
            Percent_total = round((length(which(unique(sps) %in% sp.work))/length(sp.work))*100,2))
obs.total2 <- traitslong %>%
  summarise(trait = "Total",
            N_sps = length(unique(sps)),
            N_obs = length(sps),
            Mean_obs_sps = round(N_obs/N_sps,2),
            Percent_total = round((length(which(unique(sps) %in% sp.work))/length(sp.work))*100,2))
obs.total <- rbind(obs.total, obs.total2)
DT::datatable(obs.total)
 

# N species per trait (for species in the work data)
obs.total <- traitslong[which(traitslong$sps %in% sp.work),] %>%
  group_by(trait) %>%
  summarise(N_sps = length(unique(sps)),
            N_obs = length(sps),
            Mean_obs_sps = round(N_obs/N_sps,2),
            Percent_total = round((length(which(unique(sps) %in% sp.work))/length(sp.work))*100,2))
obs.total2 <- traitslong[which(traitslong$sps %in% sp.work),] %>%
  summarise(trait = "Total",
            N_sps = length(unique(sps)),
            N_obs = length(sps),
            Mean_obs_sps = round(N_obs/N_sps,2),
            Percent_total = round((length(which(unique(sps) %in% sp.work))/length(sp.work))*100,2))
obs.total <- rbind(obs.total, obs.total2)
DT::datatable(obs.total) 

# give NA if trait value == 0
traits[traits==0] <- NA

```
We merge trait data from BIEN and TRY, giving a total of `r length(sp.traits)` species.

## Traits correlations
```{r traits_merge, echo=FALSE, message=FALSE, warning=FALSE, fig.height=6, fig.width=6}

# correlation among traits
res <- cor(traits[,-1], use = "complete.obs")
rownames(res) <- colnames(res) <- names(traits[,-1])
corrplot.mixed(res, lower = "number", upper = "circle", order = "hclust", 
               tl.col = "black", tl.srt = 45)

```

# Species in common across datasets
## Before resolving species names
```{r sp_in_common, eval=FALSE, message=FALSE, warning=FALSE}

Ns <- list(Range = sp.ranges, Phylogeny = sp.phy, Trait = sp.traits)
names(Ns) <- c(paste("Range\n(", length(sp.ranges),")",sep = ""), 
               paste("Phylogeny\n(", length(sp.phy),")",sep = ""), 
               paste("Trait\n(", length(sp.traits),")",sep = ""))

ggvenn(Ns, stroke_size = 0.5, set_name_size = 4)


```

There were `r length(Reduce(intersect, list(sp.ranges,sp.phy,sp.traits)))` species in common among the three datasets (ranges, phylogeny and traits).   

## Resolve species names
Resolve names based on the World Flora Online (WFO) taxonomic backbone.
The user needs to first download a static copy of the Taxonomic Backbone data from http://www.worldfloraonline.org/downloadData. 
We also removed taxons containing 'sp.', 'subgen.', 'sect.' and 'cf.' (N=)
```{r res_sps, eval=FALSE, message=FALSE, warning=FALSE}
# load WFO backbone
WFO <- data.table::fread("R/Cluster/Resolve_sp_names/WFO_Backbone_Plants/classification.txt", encoding="UTF-8")

# all names
Ori = gsub("_"," ",unique(c(sp.ranges,sp.phy,sp.traits))) 
length(unique(Ori)) # 153936

# Remove plant names containing 'sp.', 'subsp.', 'subgen.', 'sect.' and 'cf.'
Ori <- Ori[-grep('sp\\.',Ori)]
Ori <- Ori[-grep('subgen\\.',Ori)]
Ori <- Ori[-grep('sect\\.',Ori)]
Ori <- Ori[-grep('cf\\.',Ori)] 
length(unique(Ori)) # 153894

# Resolve species names using my own code
DevOlve <- function(x, db){
  res <- db[which(db$scientificName == x),]
  if(any(res$taxonomicStatus=="Accepted")){
    res <- res[which(res$taxonomicStatus=="Accepted"),]
  }
  else{
    if(any(res$taxonomicStatus=="Synonym")){
      id. <- res$acceptedNameUsageID
      res <- db[which(db$taxonID %in% id.),]
      res <- res[which(res$taxonomicStatus=="Accepted"),]
    }
  }
  if(nrow(res)>1){
    res <- res[-which(res$scientificNameID==""),]
    test <- res$specificEpithet == strsplit(x," ")[[1]][2]
    if(any(test)){
      res <- res[which(test),]
    } else {
      res <- res[1,]
    }
  }
  if(!nrow(res)==0){
    newname <- paste(res$genus, res$specificEpithet)
    if(any(duplicated(newname))){
      newname <- newname[-which(duplicated(newname))]
    }
    return(newname)
  }
}

cl <- makeCluster(detectCores()-1)
clusterExport(cl, c("Ori", "DevOlve", "WFO"))

newnames <- pbsapply(Ori, function(sp) DevOlve(x = sp, db = WFO),
                     cl = cl)
stopCluster(cl)
# took 27m 56s

unlistnewnames <- newnames
unlistnewnames[sapply(unlistnewnames, is.null)] <- NA

length(Ori)
length(newnames)
length(unlistnewnames)
length(unlist(unlistnewnames))

All.sps <- data.frame(Ori, newnames = unlist(unlistnewnames))

#########################
# Use “fuzzy matching” method from the R package "World Flora Online" for name identication.
# We used a maximum fuzzy distance of three to avoid incorrect species name matches.
tolook <- which(is.na(All.sps$newnames))
lookup <- All.sps$Ori[tolook]
# Divide species list into 200 chunks
spchunks <- split(lookup, sort(rep_len(1:200, length(lookup))))
# N species in each chunck
# head(lapply(spchunks, length))
# View(All.sps[tolook,])

cl <- makeCluster(detectCores()-2)
clusterExport(cl, c("spchunks", "WFO.match", "WFO"))

newnames2 <- pblapply(spchunks,
                     function(sp)
                       WFO.match(spec.data = data.frame(spec.name=sp), WFO.data = WFO),
                     cl = cl)
stopCluster(cl)
# took 03h 20m 47s

newnames2 <- lapply(newnames2, WFO.one)
newnames2 <- do.call("rbind", newnames2)

# Add new look ups to sp list
All.sps$newnames[tolook] <- paste(newnames2$genus, newnames2$specificEpithet)
View(All.sps[tolook,])
# Keep original species name for fuzzy matches greater than 3
rem <- newnames2$spec.name.ORIG[which(newnames2$Fuzzy.dist>3)]
All.sps$newnames[which(All.sps$newnames %in% rem)] <- All.sps$Ori[which(All.sps$newnames %in% rem)]
# Keep original species name for species not found
All.sps$newnames[which(All.sps$newnames=="")] <- All.sps$Ori[which(All.sps$newnames=="")]
All.sps$newnames[which(All.sps$newnames=="NULL")] <- All.sps$Ori[which(All.sps$newnames=="NULL")]
# This is the final species list
any(is.na(All.sps$newnames))
any(is.null(All.sps$newnames))
any(All.sps$newnames=="")

rownames(All.sps) <- All.sps$Ori

```

## After resolving species names
```{r sp_in_common_after, eval=FALSE, message=FALSE, warning=FALSE}

sp.ranges2 <- na.omit(unique(All.sps$newnames[which(All.sps$Ori %in% sp.ranges)]))
sp.phy2 <- na.omit(unique(All.sps$newnames[which(All.sps$Ori %in% sp.phy)]))
sp.traits2 <- na.omit(unique(All.sps$newnames[which(All.sps$Ori %in% sp.traits)]))

Ns <- list(Range = sp.ranges2, Phylogeny = sp.phy2, Trait = sp.traits2)
names(Ns) <- c(paste("Range\n(", length(sp.ranges2),")",sep = ""), 
               paste("Phylogeny\n(", length(sp.phy2),")",sep = ""), 
               paste("Trait\n(", length(sp.traits2),")",sep = ""))

ggvenn(Ns, stroke_size = 0.5, set_name_size = 4)

spincommon1 <- na.omit(unlist(Reduce(intersect, list(sp.ranges,sp.phy,sp.traits))))
spincommon2 <- na.omit(unlist(Reduce(intersect, list(sp.ranges2,sp.phy2,sp.traits2))))

# fix errors for sps in common
spincommon2 <- gsub(" ","_",spincommon2)
spincommon2 <- spincommon2[-which(spincommon2=="_")]
spincommon2 <- spincommon2[-which(spincommon2=="NA_NA")]

```

Resolving species resulted in an addition of 409 species (Before = `r length(spincommon1)`, After = `r length(spincommon2)`) in commom among the three datasets (range, traits and phy). 

## Subset datasets for species in common
I used species with at least one trait record.
```{r sp_in_common_sub, eval=FALSE, message=FALSE, warning=FALSE}

# subset phy
newphynames <- All.sps[sp.phy,]$newnames

# Fix names
newphy <- pblapply(phy, function(x) { 
  x$tip.label <- newphynames 
  x })

ncores <- detectCores()-2
cl <- makeCluster(ncores)
clusterExport(cl, c("newphy", "spincommon2", "keep.tip"))
newphy <- pblapply(newphy, function(x) keep.tip(x, spincommon2),
                   cl = cl)
stopCluster(cl)
                   
# subset traits
newtraitsnames <- All.sps[sp.traits,]$newnames
newtraits <- traits
newtraits$sps <- newtraitsnames
newtraits <- newtraits[which(newtraitsnames %in% spincommon2),]
newtraits <- newtraits[-which(duplicated(newtraits$sps)),]

# subset range data
newspnames <- All.sps[plants$sps,]$newnames
newplants <- plants
newplants$sps <- newspnames
newplants <- newplants[which(newspnames %in% spincommon2),]

# test
length(newphy[[1]]$tip.label)
length(newtraits$sps)
length(unique(newplants$sps))

```

# Imput missing traits
Because many taxa were missing some observations of certain variables, two phylogenetic gap‐filling approaches were used to estimate missing values. First, imputation was performed for a fraction of taxa that were present in the occurrence data and the phylogenetic tree, but not present in the trait data (`r length(which(!sp.traits %in% sp.work))/length(sp.work)` species).

To estimate trait values for species lacking such data, we ran phylogenetic imputation using the R-package Rphylopars (Goolsby et al. 2017) with a Brownian motion evolutionary model and took the median value from the 100 phylogenetic trees.
```{r Rphylopars, eval=FALSE, message=FALSE, warning=FALSE}

# 100 imputations >> one per phylogeny
# I was havind issues trying to run phylopars on this dataset:
# Error in tp(L = X, R = R, Rmat = as.matrix(Rmat), mL = ncol(X), mR = 1,  : 
# Not compatible with requested type: [type=character; target=double].
# Googling this error, I found this issue at the phylopars github >> https://github.com/ericgoolsby/Rphylopars/issues/21
# The workaround for the error follows:
# DO NOT RUN (JUST AN EXAMPLE)
# tmp_tree <- newphy[[1]]
# tmp_tree$edge.length[which(tmp_tree$edge.length<0.0001)]=0.0001

# save files for oppening/running in the cluster
saveRDS(newphy, "R/Cluster/Imputation/phy_plants.RDS")
saveRDS(newtraits, "R/Cluster/Imputation/traits_plants.RDS")

# Run in cluster

# Load imputed files
traitsmean <- readRDS("R/Cluster/Imputation/traitsmean.RDS")

# compare summary values of imputed vs raw traits
summary(newtraits[,-1])
summary(exp(traitsmean)) # exponention because traits were log-transformed for imputation. Exp transforms log back to normal values.

# correlation among traits
cor1 <- cor(data.frame(newtraits[,-1]), use = "complete.obs")
cor2 <- cor(exp(traitsmean))
cor1
cor2
cor1 - cor2

```

# Save data
```{r saveData, echo=TRUE, message=FALSE, warning=FALSE}

saveRDS(newplants, "Data/WorkData/plants.RDS")
saveRDS(traitsmean, "Data/WorkData/traitsmean.RDS")
saveRDS(newphy, "Data/WorkData/newphy.RDS")

```

# Save
```{r save, echo=TRUE, message=FALSE, warning=FALSE}
save.image("R/FixData.RData")
```